---
title: "Charts in RMarkdown"
author: "Michael C. Samuel, DrPH"
output: 
  html_document:
    theme: sandstone
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, fig.width = 9, fig.height = 7)
```

## About R Markdown

R Markdown is a file format for making dynamic documents with R. An R Markdown document is written in markdown (an easy-to-write plain text format) and contains chunks of embedded R code. R Markdown supports dozens of static and dynamic output formats including HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button, a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

```{r, echo=FALSE}

# Load packages
library(dplyr) # core "tidyverse" package for data manipulation
library(ggplot2) # data visualization package
library(cowplot) # ggplot add-on. Allows you to combine multiple ggplot objects into one figure
library(stringr) # package needed for character string manipulation (e.g. for wrapping text)
library(tigris) # package for reading "tiger" "shape files" for making maps
options(tigris_class = "sf") # causes tigris to read files in the "simple features" format

```

## Specify county

The data in the charts below are for the entire state. County-level data are also available by changing the **select_county** parameter below.
```{r}
select_county <- 'California'

# COVID-19 Race/Ethnicity data, filtered on the county value assigned to "select_county"
covid_re_county <- readRDS('covid_re_data.RDS') %>%
  filter(county == select_county)
```

```{r, echo = FALSE}


# California COVID-19 data 
covid_data <- readRDS('covid_data.RDS')

# Data for mapping counties
county_map <- readRDS('county_map.RDS')

# County Population Data
county_pop <- readRDS('county_population.RDS')
    
```


## COVID-19 30-Day Case, Death, and Test Rates by R/E in `r select_county`
```{r, echo=FALSE}

# Create plot
ggplot(data = covid_re_county, aes(x = demographic_set_category, y = metric_value_per_100k)) + 
  geom_bar(stat = 'identity', fill = "darkblue", colour = 'black') +
  geom_text(aes(label = paste0("N = ", scales::comma(metric_value, accuracy = 1))), vjust = 1, color = 'white') +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  labs(title = paste0("COVID-19 Rates by Race/Ethnicity in ", select_county, ", Past 30 days"), 
       x = "Race/Ethnicity", 
       y = "Rate per 100k Population") +
  facet_grid(rows = vars(metric), scales = "free") + # Forms a matrix of panels defined by faceting variable (in this case, by metric variable)
  theme_bw() + 
  theme(plot.title = element_text(face = "bold", colour = "darkblue", size = 16), 
        axis.title = element_text(face = "bold", size = 14),
        axis.text = element_text(size = 12), 
        strip.text = element_text(face = "bold", size = 12)) 

```


## Trend of COVID-19 Cases in `r select_county`
```{r, echo=FALSE}

# Filter on selected county
covid_data_county <- covid_data %>%
  filter(county == select_county)

# Create plot
ggplot(data = covid_data_county, aes(x = year_month, y = Cases)) +
  geom_line(size = 1, colour = "darkblue") +
  labs(title = paste0("Trend of COVID-19 Cases in ", select_county), 
       x = "Year-Month", 
       y = "Number of COVID-19 Cases") +
  theme_bw() + 
  theme(plot.title = element_text(face = "bold", colour = "darkblue", size = 16), 
        axis.title = element_text(face = "bold", size = 14),
        axis.text = element_text(size = 12))

```

## Trend of COVID-19 Deaths in `r select_county`
```{r, echo=FALSE}

# Filter on selected county
covid_data_county <- covid_data %>%
  filter(county == select_county)

# Create plot
ggplot(data = covid_data_county, aes(x = year_month, y = Deaths)) +
  geom_line(size = 1, colour = "darkblue") +
  labs(title = paste0("Trend of COVID-19 Deaths in ", select_county), 
       x = "Year-Month", 
       y = "Number of COVID-19 Deaths") +
  theme_bw() + 
  theme(plot.title = element_text(face = "bold", colour = "darkblue", size = 16), 
        axis.title = element_text(face = "bold", size = 14),
        axis.text = element_text(size = 12))

```


## County Map - COVID-19 Case and Death Rates
```{r, echo=FALSE}

# Sums up total # of cases and deaths within each county, and calculates case and death rates
covid_data_county <- covid_data %>%
  group_by(county) %>%
  summarise(Cases = sum(Cases), 
            Deaths = sum(Deaths)) %>%
  left_join(county_pop, by = "county") %>%
  mutate(`Case Rate` = 100000 * Cases / population, 
         `Death Rate` = 100000 * Deaths / population)

# Merge data above, and data for mapping
map_df <- county_map %>%
  left_join(covid_data_county, by = "county") 

# Create county map - case rates
case_rate_map <- ggplot(map_df) +
  geom_sf(aes(fill = `Case Rate`), size = 0.5, col = "black") +
  theme_void()

# Create county map - death rates
death_rate_map <- ggplot(map_df) +
  geom_sf(aes(fill = `Death Rate`), size = 0.5, col = "black") +
  theme_void()

# Plot case rate and death rate maps onto 1 figure
plot_grid(case_rate_map, NULL, death_rate_map, nrow = 1, rel_widths = c(0.45, 0.10, 0.45))

```
